---
title: "Reverse Engineering Project"
author: "Laura Charleston, Brennan Stewart, Nur Yavuz"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [Comic Books Are Still Made By Men, For Men and About MEN](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

marvel_data <- read_csv("marvel-wikia-data.csv")


dc_data <- read_csv("dc-wikia-data.csv")



# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 

#dc_data <- dc_data |> mutate(FIRST APPEARANCE = strftime(`FIRST APPEARANCE`, mdy(date))) |>show_col_types = FALSE



  

```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**:

    Women were underrepresented among characters with at least one appearance, and females only made up 29.3% of the DC character list and 24.7% of the Marvel roster.

-   **Analysis summary**: The results of this analysis show that both Marvel and DC have just under 30% of all their characters being female, according to the data pulled from each entity's respective Wikia page. The R analysis is in alignment with the results explained in the article.

```{r}
# Put code to reverse engineer sentence here

#dc_data <- dc_data |> group_by(SEX, APPEARANCES) |> mutate(pct_representation = Female Characters/`FIRST APPEARANCES`)*100

#dc_data <- dc_data|> mutate(percent_female = (100*n()/NROW ("Female Characters")))|> arrange (percent_female)
 

dc_data |>
  group_by(SEX)|>
  summarize(count = n())|>
  mutate(pct_female = count/6896*100)|>
  arrange(desc(pct_female))
  
marvel_data |>
  group_by(SEX)|>
  summarize(count = n())|>
  mutate(pct_female = count/16376*100)|>
    arrange(desc(pct_female))
  

# Display results of code below this codeblock



```

### Sentence 2

-   **Sentence text**: Women and men were almost exactly equally likely to have a secret identity in the Marvel universe (49.4 percent of males and 49.5 percent of females), while in the DC universe 51 percent of males had a secret identity and 45 percent of females did. 
-   **Analysis summary**: Our results are different from what the article says. From this computation, only 38% of all Marvel characters seem to have a secret identity. 27.5% of that group represents male characters, while only 9% of that group represents female characters. Looking at DC, just over 36% of superheroes have secret identities, with 9% being female and 25% being male.

```{r}
# Put code to reverse engineer sentence here

marvel_data |>
  filter(str_detect(ID, "Secret Identity")) |>
  group_by(ID, SEX) |>
  summarise(count = n()) |>
  mutate(pct_fem_secret_ID = count/16376*100)

dc_data |>
filter(str_detect(ID, "Secret Identity")) |>
  group_by(ID, SEX) |>
  summarise(count = n()) |>
  mutate(pct_fem_secret_ID = count/6896*100)
# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: Ms. Marvel is one of just 3,599 female characters in Marvel’s universe. To arrive at that number, I pulled data from the Marvel and DC Wikia databases.
-   **Analysis summary**: The finding that we discovered was 3837 female characters in Marvel's universe, which is 238 more characters than listed in the article. This may be because we are working with a more updated data set that has added more female characters since the article was published. 

```{r}
# Put code to reverse engineer sentence here

marvel_data |>
  group_by(SEX)|>
  summarize(count = n())

# Display results of code below this codeblock
  Female: 3837

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-
